---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```


## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```
## How to use the select function

Choose some columns and not others
```{r}
select(litters_df, group, litter_number)
```

This code here lets you select a range of columns 
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

You can also get rid of columns
```{r}
select(litters_df, -litter_number)

```

Renaming columns-- put the new name first 

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```
You can also use the select function for particular columns you want to keep 
Type this into the console and see objects exported from other packages above dplyr  `?select_helpers` 
```{r}
select(litters_df, starts_with("gd"))
```
This code lets you put litter number as the first column and move the others 
```{r}
select(litters_df, litter_number, everything())
```
## You can also change the rows using the filter function

Here, you can filter a column for specific values

```{r}
filter(litters_df, gd0_weight <22)
```
If you wanted to find the values for a particular column, you can filter for that value

```{r}
filter(litters_df, gd_of_birth == 20)
```
You can also search for an item that is NOT equal to that value
```{r}
filter(litters_df, gd_of_birth != 20)
```
You can search for two specifications at the same time as well 

```{r}
filter(litters_df, gd0_weight>=22, gd_of_birth == 20)
```
You can search for particular group types

```{r}
filter(litters_df, group == "Mod8")
```
I cna also search for 2 groups 

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

