---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```


## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```
## How to use the select function

Choose some columns and not others
```{r}
select(litters_df, group, litter_number)
```

This code here lets you select a range of columns 
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

You can also get rid of columns
```{r}
select(litters_df, -litter_number)

```

Renaming columns-- put the new name first 

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```
You can also use the select function for particular columns you want to keep 
Type this into the console and see objects exported from other packages above dplyr  `?select_helpers` 
```{r}
select(litters_df, starts_with("gd"))
```
This code lets you put litter number as the first column and move the others 
```{r}
select(litters_df, litter_number, everything())
```
## You can also change the rows using the filter function

Here, you can filter a column for specific values

```{r}
filter(litters_df, gd0_weight <22)
```
If you wanted to find the values for a particular column, you can filter for that value

```{r}
filter(litters_df, gd_of_birth == 20)
```
You can also search for an item that is NOT equal to that value
```{r}
filter(litters_df, gd_of_birth != 20)
```
You can search for two specifications at the same time as well 

```{r}
filter(litters_df, gd0_weight>=22, gd_of_birth == 20)
```
You can search for particular group types

```{r}
filter(litters_df, group == "Mod8")
```
I cna also search for 2 groups 

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```
There's also the mutate function that allows you to modify an existing variable or create a new variable, however this code didn't work because there are non-numeric values in the set
```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
```
You can also arrange the data set based on the values in a particular column so it's in lowest to highest etc. 

```{r}
arrange(litters_df, pups_born_alive)
```
You can also arrange so that within the order for one column, a second column's values can be in order. Just enter the columns in the order that you want them arranged.So compared to the 2 values with 3, the first value will be the lowest gd0_weight.  
```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```
## The Pipe function 
My understanding is that it's a sreies of iterative functions. I want to create the edited file litters_data_selected. 
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_data_selected
```
I REALLY like this drop_na function because it will exclude rows where a value for a particular column is gone. In this case, I will check values in the gd0_weight column 
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_with_only_complete_value = drop_na(litters_data_selected, gd0_weight)
litters_with_only_complete_value
```
If you drop the specification for a column, you can get rid of all missing data where an Na is entered

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_with_only_complete_value = drop_na(litters_data_selected)
litters_with_only_complete_value
```
Here's the pipe operator. It basically prevents you from having to retype the argument for the next function because it will automatically use it. The pipe function means do this and then do this. I'm telling it to do the same code as above.

```{r}
litters_df = read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  drop_na()
litters_df
```

