---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data

Read in the litters dataset. The janitor function cleans up the names so there are not extra spaces and characters. You can compare the file after janitors to the before file.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

```
## Take a look at the data

Printing in the console. 
```{r}
litters_df
```

## I like the skimr function because it can tell you if there are missing values for a given variable (look at the 4*8)
Here a value for gd0_weight was not entered for 15 litter samples. The histogram also can tell you if there's a bimodal distrbution of data. 
```{r}
skimr::skim(litters_df)
```
## Options to read_csv
 
If you want to learn more about the file without perusing it, type in the console `?read_csv()` for more information. 

## Other file formats

Read in an excel file. 

```{r}
library(readxl)
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df

```

Excel files also tend to have multiple sheets, so if you need help reading those, enter this in the console `?read_excel` to get more information. There you can learn how to read in only part of the file. Example

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range= "A1:F7")
mlb_df
```
Read in a SAS file. 

```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```
## Comparison with Base R
Basically don't enter a file code as read.csv instead of read_csv. It changes the names of the columns, which can be confusing.
