---
title: "Tidy Data"
output: github_document
---

```{r setup}
library(tidyverse)
```


## Pivot_longer function
This will tell you how to make your table more vertical instead of horizontal 
Load the PULSE data. Since it's a SAS file, you have to use the HAVEN package

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
pulse_data
```
##You can also go from wide format to long format--- 
This will list the variable bdi_score_bl (baseline) to bdi_score_12m (12 months) in a vertical format. That's the pivot function. The names_prefix tells the code that everything in that column is a bdi_score, so it will only output the time frames for it.I also changed those columns to just visit in general. 

```{r}
pulse_data_tidy =
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi"
    )
pulse_data_tidy

```
You can combine the above code into one! 

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )
pulse_data
```
## The RELOCATE function changes orders of columns. 
And then if you want the first 2 columns to be ID then VISIT, use the relocate function.

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id,visit) 
  
pulse_data
```
## How to have comparable values in a column
And finally, let's say I want the visit column to all have comparable values, so 00, 01, 06, 12m, I can use the mutate and recode functions. 

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m, 
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id,visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
pulse_data
```
BEAUTIFUL!! 

## Pivor_Wider is also a function 

Let's make up some data into a table! 
```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"), 
    time = c("pre", "post", "pre", "post"), 
    mean = c(4,8,3.5, 4)
  )
analysis_result
```
```{r}
analysis_result %>% 
  pivot_wider(
    names_from = "time",
    values_from = "mean"
  )
```
## Binding rows 

Using the LotR (Lord of the Ring) data. 

First step: import each table. This data set has 3 separate tables in the file. So first I have to tell it the cells that have info for that table. I added the mutate function to tell it to add a movie column and label these as fellowship ring and other movie names, so I know which movie it came from (which was listed in the excel sheet but with merged cells). 

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")
fellowship_ring

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")
two_towers

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")
return_king

```

Once I have all of the individual tables, I can bind them. 

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king)

lotr_tidy
```
Now to clean it up visually. The first argument I put in relocate comes first. 

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  relocate(movie)
lotr_tidy
```
Now to combine male and female as sex

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  relocate(movie) %>% 
  pivot_longer(
    male:female , 
    names_to = "gender", 
    values_to = "words"
  )
lotr_tidy
```

